MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads

### 2017-03-06 mlppi-comsis17/paper.tex
Intro updated on MLPPI paper

To optimize the performance of analytical processing, 
tables and materialized views in the Teradata DBMS are hash-distributed based 
on a user-specified column or set of columns called {\em primary index}. 
Each virtual processor, a unit of parallelism called {\em AMP} in 
the \hbox{Teradata} \hbox{DBMS}, 
receives a subset of the data and stores it in hash order.

Users typically choose primary index fields in the DBMS, 
so that the data is evenly distributed among the AMPs. 
The primary index fields are also chosen to reflect join fields 
in workloads to accomplish cheaper local joins that do not require 
data shuffling.

PPI (Partitioned Primary Index)~\cite{sinclair:ppi} is an optional horizontal 
partitioning scheme applied locally on each AMP's data. 
Note that in other database products, this type of partitioning would 
most likely to be called {\em clustering}. 
In the rest article we use the term \hbox{{\em partitioning}} only, not clustering, to avoid confusion. 
Database \hbox{administrators} (DBAs) usually choose the columns for PPI, 
based on join fields and single table \hbox{predicates} to optimize the queries included 
in the workload. 
The PPI columns are used to physically cluster data with the same values 
together in contiguous data blocks. 
This allows ``partition elimination'' in scans and joins for performance improvement.
PPI can be specified as single or multiple (or nested) levels. 
This type of partitioning scheme is known as 
Multi-Level PPI ({\em MLPPI)}~\cite{klindt09mlppi}.

By allowing non-qualified partitions to be eliminated, MLPPI can reduce 
significantly the amount of data to be scanned to answer a query. 
But a large number of partitions can create significant overhead, 
particularly in joins and table maintenance operations 
such as inserts and deletes, so that the selection of partitions can be usually a balancing act.

Figure~\ref{fig:exam} exemplifies an MLPPI table with potential partitioning schemes. 
This example is a subset of the {\tt LINEORDER} table from the Star Schema Benchmark (SSB)~\cite{oneil:ssb}. 
Note that the full table with modified data types was actually used in our experiments.

@@

In the definition of {\tt LINEORDER} in the figure, the primary index 
is {\tt LO\_ORDERKEY}. 
The \hbox{primary} index \hbox{dictates} the AMP on which a row will be located, 
while the partitioning of data will be dictated by the values and 
ranges associated with {\tt LO\_DISCOUNT} and {\tt LO\_QUANTITY}. 
{\tt LO\_DISCOUNT}, for example, has two ranges for \hbox{values} 
greater than or equal to 7 and one ({\tt NO CASE OR UNKNOWN}) 
for all the other values, while \linebreak {\tt LO\_QUANTITY} has four ranges. 
As a result, the relational table will have a total of 2$\times$4 = 8 partitions as follows.

\begin{center}
\begin{tabular}{|c|l|}\hline 
Partition & Condition \\ \hline
1		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY < 25}} \\ \hline
2		& {{\tt LO\_DISCOUNT >= 7 \&\& 25 <= LO\_QUANTITY <= 30}} \\ \hline
3		& {{\tt LO\_DISCOUNT >= 7 \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
4		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY no case}}\\ \hline
5       & {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY < 25}} \\ \hline 
6		& {{\tt LO\_DISCOUNT no case \&\& 25 <= LO\_QUANTITY < 30}} \\\hline								  	        
7		& {{\tt LO\_DISCOUNT no case \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
8		& {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY no case}} \\ \hline
\end{tabular}
\end{center}

Whatever partition set is chosen, 
the selection must be semantically correct; namely, 
the mapping of rows to partitions must be an injection. 
In other words, the constraints must form a covering of 
the entire range, so that a row will belong to exactly one and one partition. 
The Teradata optimizer then applies \hbox{partition} elimination for queries 
that specify conditions on {\tt LO\_DISCOUNT} and/or {\tt LO\_QUANTITY}. 
For \hbox{example}, only partitions 1 and 2 are needed for the query 
``{\tt SELECT * FROM LINEORDER WHERE LO\_DISCOUNT >= 7 AND} {\tt LO\_QUANTITY <= 30}.'' 
Similarly, partitions 1 and 5 are sufficient to answer the query 
``{\tt SELECT * FROM LINEORDER WHERE LO\_QUANTITY < 25}.'' 
\linebreak The respective sizes of the partitions are a factor of the data distribution. 
To see MLPPI in greater detail, check our references~\cite{sinclair:ppi,klindt09mlppi}.

Next, using the full version of {\tt LINEORDER} as defined in SSB~\cite{oneil:ssb} 
we provide examples of partitioning schemes that 
a DBA may define based on a small query set. 
Consider a query set $Q$ consisting of two queries $q$1 and $q$2 from the SSB set~\cite{oneil:ssb}, as shown below.

\begin{center}
{\small
\begin{tabular}{rl}
$q$1:	& {\tt SELECT SUM(l.LO$\_$EXTENDEDPRICE*l.LO\_DISCOUNT)} \\ 
		& {\tt FROM LINEORDER l, DDATE d} \\
		& {\tt WHERE l.LO\_ORDERDATE = d.D\_DATEKEY} \\
		& {\tt AND d.D\_YEAR = `1993'} \\
		& {\tt AND l.LO\_DISCOUNT IN (1, 4, 5)} \\
        & {\tt AND l.LO\_QUANTITY <= 30} \\ 
$q$2:	& {\tt SELECT c.C\_NATION, SUM(l.LO\_REVENUE)} \\ 
		& {\tt FROM CUSTOMER c, LINEORDER l} \\
		& {\tt WHERE l.LO\_CUSTKEY = c.LO\_CUSTKEY} \\
		& {\tt AND c.C\_REGION=`EUROPE'} \\
		& {\tt AND l.LO\_DISCOUNT >= 7} \\
		& {\tt AND l.LO\_QUANTITY >= 25 AND l.LO\_QUANTITY <= 35} \\
		& {\tt GROUP BY c.C\_NATION} \\
		& {\tt ORDER BY revenue desc} \\
\end{tabular}
}
\end{center}

Let's now focus on the {\tt LINEORDER} fact table and 
the predicates involving {\tt LINEORDER} fields only. 
There are five constraints identified by the query number 
and the sequence number of the constraint in each query, as illustrated below.
\vspace{-0.03in}	
\begin{center}
\begin{tabular}{|c|l|}\hline 
$q$1.1 & {\tt LO\_DISCOUNT IN (1,4,5)} \\ \hline
$q$1.2 & {\tt LO\_QUANTITY <= 30} \\ \hline
$q$2.1 & {\tt LO\_DISCOUNT >= 7} \\ \hline
$q$2.2 & {\tt LO\_QUANTITY >= 25}\\ \hline
$q$2.3 & {\tt LO\_QUANTITY <= 35} \\ \hline 
\end{tabular}
\end{center}
%\vspace{-0.05in}	
At this point the DBA needs to consider options 
based only on two fields and the five constraints. 
There are many possibilities from a {\em fine-grained} partition set 
to a loser definition.

Considering for the time being the column {\tt LO\_DISCOUNT}, 
one solution is to identify each value 
for the {\tt IN} predicate in $q$1.1 
and use $q$2.1 as is. This yields the following partitioning expression 
for {\tt LO\_DISCOUNT}:

\begin{center}
{\footnotesize
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 1,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 4,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
}
\end{center}
The above expression minimizes the size of the partitions 
by focusing exactly on the values required to satisfy the constraints, 
but creates a large number of small partitions.

Another possibility is to look at the maximum and minimum 
values in the {\tt IN} set and to build an {\tt AND} clause equivalent to 
a between clause yielding the following:
\begin{center}
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 1 AND LO\_DISCOUNT <= 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
\end{center}

@@

Another possibility is to use the partitioning associated with {\tt LO\_DISCOUNT} shown 
in the beginning of this section with only two partitions.

Similar considerations apply to the partitioning associated with {\tt LO\_QUANTITY} 
this time with ranges and covering problems difficult to deal with. 
The resulting partitioning scheme for the table includes both {\tt LO\_DISCOUNT}
and {\tt LO\_QUANTITY}, so that the number of possible combinations is the product of 
the potential combinations for each field. 
Also, there are two queries, and one partitioning scheme may be better for one query or the other. 
As a result, the DBA will be faced with a daunting combinatorial search problem 
and no clear basis to decide on which combination is the best. 
This state of affairs begs for a {\em tool} to assist the DBA.

In the sequel, we introduce an automated tool, called {\em MLPPI wizard}. 
Indeed, the \hbox{wizard} is the first physical database design tool 
developed at Teradata.
The tool is based on a novel technique using a greedy algorithm for 
search space enumeration. 
This tool based on a general framework allowing general expressions, 
ranges and case expressions for \hbox{partition} \hbox{definitions} 
is particularly well-suited for an MLPPI definition. \linebreak
The {\em predicate-driven} \hbox{technique} used by the tool can be applied 
to any clustering or partitioning based on simple fields and expressions or complex SQL predicates. 
The wizard also borrows the optimizer's cost model 
to prune the search space and reach a final solution.

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.5]{architecture}
\vspace{-0.1in}
\caption{The MLPPI wizard architecture\label{fig:arch}}
\end{figure*}

Figure~\ref{fig:arch} illustrates how our tool recommends 
the final MLPPI customized for a given workload consisting of 
a set of queries and corresponding weights. 
The wizard passes through a series of phases: {\em Preprocessing}, 
{\em Initial}, and {\em Optimized} Phases. 
In the preprocessing phase the tool produces the most granular partitioning (MGP) based on the predicates collected from every query.
In the initial phase the wizard refines the passed MGP 
by merging a pair of partitions with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}) 
and then yields an initial (feasible) MLPPI based on the refined partitions.
In the \hbox{optimization} phase, the tool makes a further attempt to optimize the initial MLPPI. 
We provide in greater \hbox{detail} each phase, 
using the query set $Q$ consisting of $q$1 and $q2$ in the rest of the article.

\shorten{
An initial MLPPI is produced through the top two phases reflecting practical constraints by field or partition. 
It is further optimized in the last phase.
In the preprocessing phase, 
for each query its predicate(s) is extracted via an API call (on {\tt explain}) made to the server. 
Then, the wizard produces the most granular partitioning (MGP) based on the 
predicates collected from every query. 
If the MGP has too many fields over a field count limit, 
then the wizard \hbox{incrementally} throws away a field from the MGP until the limit.

In the initial phase, the passed MGP gets reduced if it includes more partitions than a partition count limit allowed by an MLPPI. 
The refinement can be \hbox{incrementally} done by merging a pair of partitions 
with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}), 
until the limit. 
After the phase, an initial (feasible) MLPPI is produced based on the refined partitions.

In the optimization phase, the tool makes a further attempt to optimize the initial MLPPI. 
Similar to the initial phase, more partitions can be reduced by merging ones with the 
least {\em query cost} (as will be described in Section~\ref{sec:opt_phase}). 
If the ongoing solution cannot be further improved, the tool makes the final MLPPI recommendation. 
We will provide more details about each phase later, 
using the query set $Q$ consisting of $q$1 and $q2$ 
in the rest of the article.}

To the best of our knowledge, there is no existing industrial tool 
to solve the {\it {\bf multi-level partitioning}} problem except our wizard. 
The wizard is contained completely on the client side, as opposed to 
previous work~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} 
requiring optimizer code extension. 
Our tool simply uses existing APIs to simplify the queries, capture fact table 
predicates and costs, and uses these items to make a recommendation. 
This makes the wizard extensible and portable to different releases 
of the Teradata DBMS server.

>>>>

%\subsection{Background}
%\label{sec:background}

\paragraph{Background:} 
To optimize the performance of analytical processing, 
tables and materialized views in the Teradata DBMS are hash-distributed based 
on a user-specified column or set of columns called {\em primary index}. 
Each virtual processor, a unit of parallelism called {\em AMP} in 
the \hbox{Teradata} \hbox{DBMS}, 
receives a subset of the data and stores it in hash order. 
Users typically choose primary index fields in the DBMS, 
so that the data is evenly distributed among the AMPs. 
The primary index fields are also chosen to reflect join fields 
in workloads to accomplish cheaper local joins that do not require data shuffling.

Partitioned Primary Index (PPI)~\cite{sinclair:ppi} is an optional horizontal 
partitioning scheme applied locally on the data belonging to each AMP. 
In some database products, this type of partitioning is termed {\em clustering}, 
but hereafter we call the term \hbox{{\em partitioning}} only to avoid confusion.
Database \hbox{administrators} (DBAs) usually choose the columns for PPI, 
based on join fields and single table \hbox{predicates} to optimize the queries included 
in the workload. 
The PPI columns are used to physically group data with the same values 
together in contiguous data blocks. 
This grouping enables ``partition elimination'' in scans and joins for performance enhancement.
PPI can be specified as single or multiple levels. 
This type of partitioning scheme is known as 
\hbox{Multi-Level} PPI ({\em MLPPI)}~\cite{klindt09mlppi}.

By allowing non-qualified partitions to be eliminated, MLPPI can considerably 
reduce the amount of data to be scanned to answer a query. 
But a large number of partitions can create significant overhead, 
particularly in joins and table maintenance operations 
such as inserts and deletes, so that the selection of partitions can be usually a balancing act.

Figure~\ref{fig:exam} exemplifies an MLPPI table with a defined partitioning option. 
This example is a subset of the {\tt LINEORDER} table from the Star Schema Benchmark (SSB)~\cite{oneil:ssb}. 
(Note that the full table with modified data types was actually used in our experiments.)

@@

In the definition of {\tt LINEORDER} in the figure, the primary index 
is on {\tt LO\_ORDERKEY}. 
The \hbox{primary} index \hbox{dictates} the AMP on which a row will be located 
while the partitioning of data will be dictated by the values and 
ranges associated with {\tt LO\_DISCOUNT} and {\tt LO\_QUANTITY}. 
For example, {\tt LO\_DISCOUNT} has two ranges: (i)~one for \hbox{values} $\geq$ 7 
and (ii)~another for all the other values (or {\tt NO CASE OR UNKNOWN}). 
{\tt LO\_QUANTITY} has four ranges. 
Hence, the relation will have a total of 2$\times$4 = 8 partitions as follows.

\begin{center}
{\footnotesize
\begin{tabular}{|c|l|}\hline 
Partition Number & Condition \\ \hline
1		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY < 25}} \\ \hline
2		& {{\tt LO\_DISCOUNT >= 7 \&\& 25 <= LO\_QUANTITY <= 30}} \\ \hline
3		& {{\tt LO\_DISCOUNT >= 7 \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
4		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY no case}}\\ \hline
5       & {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY < 25}} \\ \hline 
6		& {{\tt LO\_DISCOUNT no case \&\& 25 <= LO\_QUANTITY < 30}} \\\hline								  	        
7		& {{\tt LO\_DISCOUNT no case \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
8		& {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY no case}} \\ \hline
\end{tabular}
}
\end{center}

Whatever partition is chosen, 
the selection must be semantically correct; namely, 
the mapping of rows to partitions must be an injection. 
In other words, the constraints must form a covering of 
the entire range, so that a row will belong to exactly one and one partition. 
The Teradata optimizer then applies \hbox{partition} elimination for queries 
that specify conditions on {\tt LO\_DISCOUNT} and/or {\tt LO\_QUANTITY}. 
For \hbox{example}, only partitions~1 and~2 are needed for the query 
``{\tt SELECT * FROM LINEORDER WHERE LO\_DISCOUNT >= 7 AND} {\tt LO\_QUANTITY <= 30}.'' 
Similarly, partitions~1 and~5 are sufficient to answer the query 
``{\tt SELECT * FROM LINEORDER WHERE LO\_QUANTITY < 25}.'' 
\linebreak The respective sizes of the partitions are a factor of the data distribution. 
To see MLPPI in greater detail, check our references~\cite{klindt09mlppi,sinclair:ppi}.

%Next, using {\tt LINEORDER} defined in the SSB we provide examples of partitioning options that 
%a DBA may define based on a small query set. 
\paragraph{Challenge:}
Consider a query set $Q$ having two queries $q$1 and $q$2 on the SSB, as shown below.
\vspace{-.05in}
\begin{center}
{\small
\begin{tabular}{rl}
$q$1:	& {\tt SELECT SUM(l.LO$\_$EXTENDEDPRICE*l.LO\_DISCOUNT)} \\ 
		& {\tt FROM LINEORDER l, DDATE d} \\
		& {\tt WHERE l.LO\_ORDERDATE = d.D\_DATEKEY} \\
		& {\tt AND d.D\_YEAR = `1993'} \\
		& {\tt AND l.LO\_DISCOUNT IN (1, 4, 5)} \\
        & {\tt AND l.LO\_QUANTITY <= 30} \\ 
$q$2:	& {\tt SELECT c.C\_NATION, SUM(l.LO\_REVENUE)} \\ 
		& {\tt FROM CUSTOMER c, LINEORDER l} \\
		& {\tt WHERE l.LO\_CUSTKEY = c.LO\_CUSTKEY} \\
		& {\tt AND c.C\_REGION=`EUROPE'} \\
		& {\tt AND l.LO\_DISCOUNT >= 7} \\
		& {\tt AND l.LO\_QUANTITY >= 25 AND l.LO\_QUANTITY <= 35} \\
		& {\tt GROUP BY c.C\_NATION} \\
		& {\tt ORDER BY revenue desc} \\
\end{tabular}
}
\vspace{-.01in}
\end{center}

A DBA may focus on the {\tt LINEORDER} fact table and 
the predicates involving {\tt LINEORDER} fields only. 
He can figure out the following five constraints, identified by the query number 
and the sequence number of predicates in each query: 
\vspace{-.05in}
\begin{center}
\begin{tabular}{cl}
$q$1.1: & {\tt LO\_DISCOUNT IN (1,4,5)} \\ 
$q$1.2: & {\tt LO\_QUANTITY <= 30} \\ 
$q$2.1: & {\tt LO\_DISCOUNT >= 7} \\ 
$q$2.2: & {\tt LO\_QUANTITY >= 25}\\ 
$q$2.3: & {\tt LO\_QUANTITY <= 35}. \\ 
\end{tabular}
\end{center}
\vspace{-.05in}
At this point the DBA needs to take into account options 
based only on two fields and the five constraints. 
There are many other possibilities from a {\em fine-grained} partition set 
to a {\em coarser-grained} definition.

Considering for the time being the column {\tt LO\_DISCOUNT}, 
one solution is to identify each value 
for the {\tt IN} predicate in $q$1.1 
and use $q$2.1 as is. This yields the following partitioning expression 
for {\tt LO\_DISCOUNT}:
\vspace{-.05in}
\begin{center}
{\footnotesize
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 1,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 4,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
}
\end{center}
The above expression minimizes the size of the partitions 
by focusing exactly on the values required to satisfy the constraints 
but creates a large number of small partitions.

Another possibility is to look at the maximum and minimum 
values in the {\tt IN} set and to build an {\tt AND} clause equivalent to 
a between clause yielding the following:
\vspace{-.1in}
\begin{center}
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 1 AND LO\_DISCOUNT <= 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
\vspace{-.1in}
\end{center}

@@

(Another possibility would be to use the partitioning associated with {\tt LO\_DISCOUNT} shown 
in the beginning of this section with only two partitions.)

Similar considerations apply to the partitioning associated with {\tt LO\_QUANTITY}. 
The resulting partitioning definition for the table includes both {\tt LO\_DISCOUNT}
and \linebreak {\tt LO\_QUANTITY}, and thus the number of possible combinations is the product of 
the potential combinations for each field.

Furthermore, given two queries one partitioning scheme may be favorable for one query but not for the other, or vice versa. 
As a result, the DBA becomes faced with a daunting combinatorial search problem 
and no clear basis to decide on which combination is the best. 
This state of affairs begs for a {\em tool} for the DBA.

%\subsection{Solution: An Automated Partitioning Tool}
%\label{sec:solution}
\paragraph{Solution:} 
To ease a burden of partitioning decision, 
we introduce a novel physical database design tool, called {\em MLPPI wizard}, 
designed to recommend an effective partitioning solution for a given workload.
%Indeed, the \hbox{wizard} is the first physical database design tool 
%developed at Teradata.
This tool uses a greedy algorithm for search space enumeration 
and is based on a general framework allowing general expressions, 
ranges and case expressions for partition definitions. 
%is particularly well-suited for an MLPPI definition. 
We emphasize that the \hbox{predicate-driven} algorithm used by the tool can be applied 
to {\em any} clustering or partitioning scheme based on simple fields and expressions or complex SQL predicates 
in an {\em arbitrary} DBMS. 
The wizard also borrows the optimizer's cost model 
to prune the search space and reach a final solution.

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.48]{architecture}
\vspace{-0.1in}
\caption{The MLPPI wizard architecture\label{fig:arch}}
\end{figure*}

Figure~\ref{fig:arch} illustrates how our tool recommends 
the final MLPPI customized for a given workload consisting of 
a set of queries and corresponding weights. 
The wizard passes through a series of phases: {\em Preprocessing}, 
{\em Initial}, and {\em Optimized} Phases. 
In the pre-processing phase the tool produces the most granular partitioning (MGP) based on the predicates collected from every query.
In the initial phase the wizard refines the passed MGP 
by merging a pair of partitions with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}) 
and then yields an initial (feasible) MLPPI based on the refined partitions.
In the \hbox{optimization} phase, the tool makes a further attempt to optimize the initial MLPPI. 
We elaborate in greater detail on each phase, 
using the running query set $Q$ consisting of $q$1 and $q2$ in the rest of the article.

%\shorten{
%An initial MLPPI is produced through the top two phases reflecting practical constraints by field or partition. 
%It is further optimized in the last phase.
%In the preprocessing phase, 
%for each query its predicate(s) is extracted via an API call (on {\tt explain}) made to the server. 
%Then, the wizard produces the most granular partitioning (MGP) based on the 
%predicates collected from every query. 
%If the MGP has too many fields over a field count limit, 
%then the wizard \hbox{incrementally} throws away a field from the MGP until the limit. 
%
%In the initial phase, the passed MGP gets reduced if it includes more partitions than a partition count limit allowed by an MLPPI. 
%The refinement can be \hbox{incrementally} done by merging a pair of partitions 
%with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}), 
%until the limit. 
%After the phase, an initial (feasible) MLPPI is produced based on the refined partitions.
%
%In the optimization phase, the tool makes a further attempt to optimize the initial MLPPI. 
%Similar to the initial phase, more partitions can be reduced by merging ones with the 
%least {\em query cost} (as will be described in Section~\ref{sec:opt_phase}). 
%If the ongoing solution cannot be further improved, the tool makes the final MLPPI recommendation. 
%We will provide more details about each phase later, 
%using the query set $Q$ consisting of $q$1 and $q2$ 
%in the rest of the article.}

To the best of our knowledge, there is few extant industrial tool 
to addressing the {\it {\bf \hbox{multi-level} partitioning}} problem, except our wizard. 
The wizard resides completely on the client side as opposed to 
several previous tools~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} 
that require optimizer code instrumentation. 
Our tool simply uses existing APIs to simplify the queries, capture fact table 
predicates and costs, and uses these items to make a recommendation. 
%This makes the wizard extensible and portable to different releases 
%of the Teradata DBMS server. 
This \hbox{loosely-coupled} approach makes the wizard extensible and portable to different releases 
of the database server.
